---
title: "Creating a UAT performance test dataset using `seqgendiff`"
author: "Alex Kalinka"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    depth: 3
    highlight: tango
    number_sections: true
    theme: spacelab
    toc: true
    toc_float: true
---

# Setup

```{r setup, include=FALSE}
set.seed(1)

options(warn=-1)
suppressMessages(library(knitr))
suppressMessages(library(tidyr))
suppressMessages(library(magrittr))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(seqgendiff))

c903 <- read.table("../data/HT29_c903.tsv", header=T, stringsAsFactors = F)[,c(1:7)] %>%
  dplyr::filter(rowSums(.[,3:6]) > 30)

```

# Summary

`seqgendiff` will be used to add signal (using 'binomial thinning') to an original dataset consisting of 4 HT29 control replicates in which there should not be any significant differences (i.e. all true negatives prior to the addition of signal). In this way, there will be a set of known true positives and true negatives allowing estimation of the false negative and false positive rates of the calling algorithm (in this case `Mageck`).

# Creating the data

* 500 genes will be randomly selected to receive signal in their gRNAs.
* Log-fold change (logFC) signal randomly sampled using `rnorm(500,0,1)`, which would result in approx. 250 genes with an expected absolute logFC > 0.8.
* All guides belonging to a given gene will receive this signal via a noisy `binomial thinning` process.

```{r}
# Randomly sample genes to receive a signal.
lfc <- rnorm(500)
names(lfc) <- sample(unique(c903$gene), 500, replace=F)

coef_mat <- as.matrix(c903 %>%
                        dplyr::select(gene,-ERS717283.plasmid) %>%
                        dplyr::mutate(gene_indicator = ifelse(gene %in% names(lfc),1,0)) %>%
                        dplyr::group_by(gene) %>%
                        dplyr::mutate(gene_indicator = ifelse(gene[1] %in% names(lfc),
                                                              rep(lfc[names(lfc)==gene[1]],n()),
                                                              rep(0,n()))) %>%
                        dplyr::ungroup() %>%
                        dplyr::select(gene_indicator))

design_mat <- matrix(c(0,0,1,1))
colnames(design_mat) <- "treatment"

# Add signal to randomly sampled genes and their gRNAs.
thout <- thin_diff(mat = as.matrix(c903[,4:7]), 
                   design_fixed = design_mat, 
                   coef_fixed = coef_mat)

c903.bthin <- data.frame(c903[,1:2], thout$mat, stringsAsFactors = F)
colnames(c903.bthin)[3:6] <- c(paste("control.",1:2,sep=""),paste("treat.",1:2,sep=""))

```

```{r}
# For the AZ_CRUK pipeline, we need a 'baseline' and 'plasmid' sample for a resistance screen.
# Here, we add baseline and plasmid columns from the venetoclax datasets.
vclax <- read.table("../data/venetoclax_d7_vs_DMSO_combined_counts.txt", 
                    header=T, stringsAsFactors = F)
c903.bthin.pipeline <- c903.bthin %>%
  mutate(YUSAplasmid = vclax$YUSAplasmid[match(sgRNA, vclax$sgRNA)],
         YUSADMSOD0 = vclax$YUSADMSOD0[match(sgRNA, vclax$sgRNA)],
         YUSAplasmid = ifelse(is.na(YUSAplasmid),10,YUSAplasmid),
         YUSADMSOD0 = ifelse(is.na(YUSADMSOD0),10,YUSADMSOD0))

```

# Algorithm tests

```{r}
# Limit of Detection (LOD) log2 fold change (for genes with a true signal).
# Genes with 0 < lfc < 0.8 will be ignored when calculating sens. and spec.
lod.lfc <- 0.8
```

## Mageck

```{r}
c903.mageck <- read.table("../data/c903-bthin-rnorm.500.symmFC.gene_summary.txt",
                          header=T, stringsAsFactors = F) %>%
  mutate(lfc_bthin = ifelse(id %in% names(lfc), lfc[match(id, names(lfc))], 0)) %>%
  rowwise() %>%
  mutate(sign.05 = (neg.fdr < 0.05 || pos.fdr < 0.05),
         sign.1 = (neg.fdr < 0.1 || pos.fdr < 0.1),
         ignore = (abs(lfc_bthin) > 0 && abs(lfc_bthin) < lod.lfc),
         true_neg = lfc_bthin==0,
         true_pos = abs(lfc_bthin)>0,
         false_pos = (sign.1 && lfc_bthin==0),
         false_neg = (!sign.1 && lfc_bthin>0)) %>%
  ungroup()

c903.mageck %>%
  filter(!ignore) %>%
  summarise(specificity = 1-sum(false_pos)/sum(true_neg),
            sensitivity = 1-sum(false_neg)/sum(true_pos))

ggplot(c903.mageck,aes(lfc_bthin,neg.lfc)) +
  geom_point() +
  geom_point(data = c903.mageck %>% filter(sign.05), aes(color="signif.")) +
  geom_abline(linetype="dashed") +
  xlab("Expected logFC") + ylab("Observed logFC") +
  ggtitle("Mageck hits at 10% FDR")

```

## DrugZ

```{r}
c903.drugz <- read.table("../data/c903-bthin-rnorm.500.symmFC--drugz.txt",
                          header=T, stringsAsFactors = F) %>%
  mutate(lfc_bthin = ifelse(GENE %in% names(lfc), lfc[match(GENE, names(lfc))], 0)) %>%
  rowwise() %>%
  mutate(sign.05 = (fdr_synth < 0.05 || fdr_supp < 0.05),
         sign.1 = (fdr_synth < 0.1 || fdr_supp < 0.1),
         ignore = (abs(lfc_bthin) > 0 && abs(lfc_bthin) < lod.lfc),
         true_neg = lfc_bthin==0,
         true_pos = abs(lfc_bthin)>0,
         false_pos = (sign.1 && lfc_bthin==0),
         false_neg = (!sign.1 && lfc_bthin>0)) %>%
  ungroup()

c903.drugz %>%
  filter(!ignore) %>%
  summarise(specificity = 1-sum(false_pos)/sum(true_neg),
            sensitivity = 1-sum(false_neg)/sum(true_pos))

ggplot(c903.drugz,aes(lfc_bthin,normZ)) +
  geom_point() +
  geom_point(data = c903.drugz %>% filter(sign.05), aes(color="signif.")) +
  geom_abline(linetype="dashed") +
  xlab("Expected logFC") + ylab("Observed normZ score") +
  ggtitle("DrugZ hits at 10% FDR")

```


